---
title: "fit_batch_clean"
output: html_document
---

```{r sourcce, message=FALSE, warning=FALSE}
setwd(dirname(getwd()))
source('paths.R', encoding="utf8" )
source( file.path(post.dir, 'diagnostics.R'), encoding="utf8" )
source( file.path(fit.dir, 'fit.R'), encoding="utf8" )
source( file.path(dataprep.dir, 'dataprep.R'), encoding="utf8" )
library(rstan)
library(ggplot2)
```

## data:

N : int, number of languages
feats_lang : data.frame, feats of languages for current family
trees.C : list, covariance matrices for language trees

## features:

- NA : noun adjective order
    0: noun-adjective, 1:adjective-noun, -: other
  rename to AN, to avoid naming conflict
    
- PN : adposition noun order
    0: prepositions, 1: adpositions, -: other
    
- ND : noun demonstrative order
    0: noun-demonstrative, 1: demonstrative-noun, -: other

- NG : noun genetive order
    0: noun-genetive, 1: genetive-noun, -: other

- NNum : noun numeral order
    0: noun-numeral, 1: numeral-noun, -: other

- VO : verb object order
    0: verb-object, 1:object-verb, -: other

- NRc : noun relative clause order
    0: noun-relative clause, 1: relative clause-noun, -: other

- VS : verb subject order
    0: verb-subject, 1: subject-verb, -: other
```{r compile models}
options(mc.cores = parallel::detectCores())
memory.limit( size=30000 )
chains <- 6
max.C <- 5

model.dep <- cmdstanr::cmdstan_model( 
  file.path( models.dir, 'CV', 'TwoTraits-binom-multiC-cholesky-nonCentered-CV-fullEval.stan' ),
  pedantic=TRUE)
model.indep <- cmdstanr::cmdstan_model( 
  file.path( models.dir, 'CV', 'TwoTraits-indep-binom-multiC-cholesky-nonCentered-CV-fullEval.stan' ),
  pedantic=TRUE)
```

```{r fit Models}
#lang.fam.names <- c( "Afro-Asiatic", "Algic", "Arawakan", "Athapaskan-Eyak-Tlingit", "Atlantic-Congo",
#                    "Austroasiatic", "Austronesian", "Cariban", "Central_Sudanic", "Chibchan", 
#                    "Dravidian", "Gunwinyguan", "Indo-European", "Mande", "Mongolic",  
#                    "Nakh-Daghestanian", "Nilotic", "Nuclear_Torricelli", "Nuclear_Trans_New_Guinea", 
#                    "Otomanguean", "Pama-Nyungan", "Pano-Tacanan", "Salishan", "Sepik", "Sino-Tibetan",
#                    "Siouan", "Sko", "Surmic", "Tai-Kadai", "Tucanoan", "Tupian", "Turkic", "Uralic",
#                    "Uto-Aztecan" )
#lang.fam.names <- c( "Arawakan", "Chibchan", "Gunwinyguan", "Nilotic", 
#                     "Pano-Tacanan", "Sko", "Surmic", "Tai-Kadai", "Tucanoan",
#                     "Uralic" )
lang.fam.names <- c( "Uto-Aztecan" )
trait.names <- c( "AN", "PN", "ND", "NG", "NNum", "VO", "NRc", "VS" )
t.pairs <- c()

trait.pairs <- vector(mode="character", length=28)
acc.dep <- vector(mode="numeric", length=28)
acc.indep <- vector(mode="numeric", length=28)
diff.1 <- vector(mode="numeric", length=28)
diff.2 <- vector(mode="numeric", length=28)
upper.50 <- vector(mode="logical", length=28)
j <- 1

for (trait.name.1 in trait.names){
  for (trait.name.2 in trait.names){
    pair = paste( trait.name.1, trait.name.2 )
    
    if ( (trait.name.1 != trait.name.2) & !(pair %in% t.pairs) ){
      
      t.pairs = c( t.pairs, pair, paste(trait.name.2, trait.name.1) )
      
      dep.mN = NA
      indep.mN = NA
      
      for ( i in 1:length(lang.fam.names) ){
        
        lang <- lang.fam.names[i]
        
        dat <- get.data.3fold( lang.name=lang, trait.name.1=trait.name.1,
                               trait.name.2=trait.name.2,
                               M=max.C, seed.N=1, p.test=3,
                               fullEval=TRUE )
        dat.1 <- list(
          N_train = dat$train.N,
          N = dat$test.N,
          x_train = dat$fold.1$train.traits,
          x = dat$fold.1$test.traits,
          unseen = dat$fold.1$unseen,
          M = max.C,
          C_train = dat$fold.1$train.C,
          C = dat$fold.1$test.C,
          dMat_train = designMatrix(dat$train.N,2),
          dMat = designMatrix(dat$test.N,2),
          # Hyperparameters
          eta = 0.8,
          lambda = 1,
          mu_z = 0,
          sigma_z = 1.5
          )
        dat.2 <- list(
          N_train = dat$train.N,
          N = dat$test.N,
          x_train = dat$fold.2$train.traits,
          x = dat$fold.2$test.traits,
          unseen = dat$fold.2$unseen,
          M = max.C,
          C_train = dat$fold.2$train.C,
          C = dat$fold.2$test.C,
          dMat_train = designMatrix(dat$train.N,2),
          dMat = designMatrix(dat$test.N,2),
          # Hyperparameters
          eta = 0.8,
          lambda = 1,
          mu_z = 0,
          sigma_z = 1.5
          )
        dat.3 <- list(
          N_train = dat$train.N,
          N = dat$test.N,
          x_train = dat$fold.3$train.traits,
          x = dat$fold.3$test.traits,
          unseen = dat$fold.3$unseen,
          M = max.C,
          C_train = dat$fold.3$train.C,
          C = dat$fold.3$test.C,
          dMat_train = designMatrix(dat$train.N,2),
          dMat = designMatrix(dat$test.N,2),
          # Hyperparameters
          eta = 0.8,
          lambda = 1,
          mu_z = 0,
          sigma_z = 1.5
          )

        print( paste("--", trait.name.1, "-", trait.name.2, "-", lang, "--") )
        
        print( "- dependent ")
        dependent <- fit.CV.3f( model=model.dep, dat.1=dat.1, dat.2=dat.2, dat.3=dat.3, indep=FALSE, 
                                lang=lang, trait.name.1=trait.name.1, trait.name.2=trait.name.2 )
        print( "- independent ")
        independent <- fit.CV.3f( model=model.indep, dat.1=dat.1, dat.2=dat.2, dat.3=dat.3, indep=TRUE, 
                                  lang=lang, trait.name.1=trait.name.1, trait.name.2=trait.name.2 )
        
        trait.pairs[j] <- pair
        acc.dep[j] <- dependent$acc
        acc.indep[j] <- independent$acc
        diff.1[j] <- dependent$acc - independent$acc
        diff.2[j] <- independent$acc - dependent$acc
        upper.50[j] <- (dependent$acc + independent$acc) > 1
        j <- j+1
      }
    }
  }
}

```
```{r plot accuracies}
unknowns <- dat$test.N - dat$train.N
chance <- 0.5 ^ unknowns
#chance <- 0.5
acc.df.a <- data.frame( trait.pairs=as.factor(trait.pairs), accuracy=acc.dep, 
                        model=rep("dependent", length(acc.dep)) )
acc.df.b <- data.frame( trait.pairs=as.factor(trait.pairs), accuracy=acc.indep,
                        model=rep("independent", length(acc.indep)) )
acc.df <- rbind(acc.df.a, acc.df.b)

ggplot(acc.df) + 
  scale_y_continuous(
    limits=c(NA, 1), breaks=seq(from=0, to=1, by=0.1)
    ) +
  geom_col(data=acc.df, aes(x=trait.pairs, y=accuracy, fill=model), position="dodge2") +
  scale_fill_manual(values = c("green4", "blue4") ) +
  geom_hline(yintercept=chance, color="red", size=1) +
  coord_flip() + labs( title=lang )
```

```{r plot diffs}
dAccdf <- data.frame(trait.pairs=as.factor(trait.pairs), dAccuracy=diff.1, upper.50=upper.50)
pos <- dAccdf[dAccdf$dAccuracy>=0,]
neg <- dAccdf[dAccdf$dAccuracy<0,]
u.5 <- dAccdf[dAccdf$upper.50==TRUE,]

ggplot(dAccdf) + 
  geom_col(data=pos, aes(x=trait.pairs, y=dAccuracy), fill="green4") +
  geom_col(data=neg, aes(x=trait.pairs, y=dAccuracy), fill="blue4") + 
  geom_hline(yintercept=0, color="red", size=1.2) +
  coord_flip() + labs( title=lang, subtitle="dependent over independent" )


dAccdf.2 <- data.frame(trait.pairs=as.factor(trait.pairs), dAccuracy=diff.2)
pos <- dAccdf.2[dAccdf.2$dAccuracy>=0,]
neg <- dAccdf.2[dAccdf.2$dAccuracy<0,]
u.5 <- dAccdf.2[dAccdf.2$upper.50==TRUE,]

ggplot(dAccdf.2) + 
  geom_col(data=pos, aes(x=trait.pairs, y=dAccuracy), fill="green4") +
  geom_col(data=neg, aes(x=trait.pairs, y=dAccuracy), fill="blue4") + 
  geom_hline(yintercept=0, color="red", size=1.2) +
  coord_flip() + labs( title=lang, subtitle="independent over dependent" )
```

```{r}
beepr::beep(8)
```



















